---
title: "Chapter 3 Homework"
author: "Maddie Thall"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(knitr)
```

**Question 1:**  
There are different sampling designs used to ensure random sampling, including
simple random sampling, systematic random sampling, stratified random sampling, 
and cluster sampling. Random sampling can happen both with and without replacement.
A researcher may choose not to sample with replacement to maximize the diversity of
samples.  

**Question 2:**  
AA: 0.85  
AS: 0.148    
SS: 0.0064  
  
**Question 3:**  

```{r}
frequency = c(8, 1, 8, 1, 1, 1, 1)
labels = c("Hispanic", "Hmong", "African American", "Japanese", "Greek", "Chinese", "Samoan")
df = data.frame(labels, frequency)
df = df %>%
  mutate(
    percent = frequency / sum(frequency) * 100,
    expected_80 = frequency / sum(frequency) * 80)
df = df %>%
  rename(
    "Ethnic Group" = labels,
    "Observed Frequency" = frequency,
    "Percent (%)" = percent,
    "Expected Count (n = 80)" = expected_80
  )
kable(df, caption = "Probability Distribution of Ethnic Groups with Expected Counts",
      digits = 2)
```


**Question 4:**
```{r}
berries_gathered = c(25, 30, 50, 65, 70, 80, 95)
intervals = c(1, 1, 1, 2, 2, 1, 1)
N = sum(intervals)
lambda = sum(berries_gathered * intervals) / N
relative_freq = dpois(berries_gathered, lambda)
absolute_freq = relative_freq * N
df_poisson = data.frame(
  Berries_Gathered = berries_gathered,
  Observed_Frequency = intervals,
  Relative_Expected = dpois(berries_gathered, lambda),
  Absolute_Expected = dpois(berries_gathered, lambda) * N
)

kable(df_poisson, caption = "Observed and Expected Poisson Frequencies for Berry Gathering", digits = 5)
```
These data do not follow a Poisson distribution because some of the expected frequencies 
are extremely small, with many frequencies close to zero. Also, the observed distribution 
is much more dispersed than expected under Poisson assumptions.  
  
  
**Question 5:**  
```{r}
mu = 5
sigma = 2

p_a = pnorm(1, mean = mu, sd = sigma)

p_b = pnorm(6, mean = mu, sd = sigma) - pnorm(3, mean = mu, sd = sigma)

p_c = pnorm(4, mean = mu, sd = sigma) - pnorm(2, mean = mu, sd = sigma)

p_d = 1 - pnorm(9, mean = mu, sd = sigma)

results = data.frame(
  Question = c("$\\le 1$ lesion",
               "3–6 lesions",
               "2–4 lesions",
               "$\\ge 9$ lesions"),
  Probability = round(c(p_a, p_b, p_c, p_d), 3)
)

kable(results, caption = "Probabilities for Lesion Counts in a Normal Distribution")

```

